Capstone Movielens Report

### 2025-12-10 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `calc_UMGY_SmoothedDay_effect`.

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UM_effect.cv.details) section of *This Description* below.
:::
:::

@@

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMG_effect.cv.details) section of *This Description* below.
:::
:::

@@

An example of using it in the [train_UMGY_effect](#func.train_UMGY_effect) function is shown above in the [Usage](#func.calc_UMGY_effect.usage) section of *This Description*:

@@

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMGY_effect.cv.details) section of *This Description* below.
:::
:::

@@

Builds the *User+Movie+Genre+Year Effect (UMGYDE) Model* based on the *Day General Effect* data structure returned by the [calc_day_general_effect](#func.calc_day_general_effect) or [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) function (both described above) and passed in the [date_general_effect](#func.calc_UMGY_SmoothedDay_effect.args.date_general_effect) argument.

@@

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_UMGY_SmoothedDay_effect` is called internally from the [train_UMGY_effect](#func.train_UMGY_effect) function described below.
:::
:::

@@

::: { #func.calc_UMGY_SmoothedDay_effect.args.date_general_effect .sidebar }
  - **date_general_effect:** Data frame representing the *Day General Effect*;

@@

An example of using it in the [train_UMGY_effect](#func.train_UMGY_effect) function is shown above in the [Usage](#func.calc_UMGY_SmoothedDay_effect.usage) section of *This Description*:

@@

The source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L3) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) script on _GitHub_.
:::
:::

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_SmoothedDay_effect.args.lambda) argument value is non-default) the *User+Movie+Genre+Year Effect (UMGYDE) Model* using the training sample of the `Movielens` dataset passed in the [train_set](#func.train_UMGY_SmoothedDay_effect.args.train_set) argument.

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_SmoothedDay_effect.cv.args.lambda) argument value is non-default) the *User+Movie+Genre+Year Effect (UMGYDE) Model* with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

Trains (and optionally *regularizes* if the [lambda](#func.regularize.train_UMGYD_effect.args.lambda) argument value is non-default) the *User+Movie+Genre+Year Effect (UMGYDE) Model* using the training sample of the `Movielens` dataset passed in the [train_set](#func.regularize.train_UMGYD_effect.args.train_set) argument.

@@

Trains and *regularizes* the *User+Movie+Genre+Year Effect (UMGYDE) Model* with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMGYD_effect.cv.args.lambda) argument for the *User+Movie+Genre+Year Effect (UMGYDE) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.

@@

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMGYD_effect.cv.details) section of *This Description* below.
:::
:::

>>>>

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UM_effect.cv.details) section of *this Description* below.
:::
:::

@@

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMG_effect.cv.details) section of *this Description* below.
:::
:::

@@

An example of using it in the [train_UMGY_effect](#func.train_UMGY_effect) function is shown above in the [Usage](#func.calc_UMGY_effect.usage) section of *this Description*:

@@

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMGY_effect.cv.details) section of *this Description* below.
:::
:::

@@

Builds the *User+Movie+Genre+Year+SmoothedDay Effect (hereafter UMGYDE for short) Model* based on the *Day General Effect* (*DG Effect* for short) data structure returned by the [calc_day_general_effect](#func.calc_day_general_effect) or [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) function (both described above) and passed in the [date_general_effect](#func.calc_UMGY_SmoothedDay_effect.args.date_general_effect) argument.

@@

```{r eval=FALSE}
  train_set |> 
    calc_day_general_effect(lambda) |>
    calc_UMGY_SmoothedDay_effect(degree, span)
```

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_UMGY_SmoothedDay_effect` is called internally from the [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect) function described below.
:::
:::

@@

::: { #func.calc_UMGY_SmoothedDay_effect.args.date_general_effect .sidebar }
  - **date_general_effect:** Dataframe object representing the *DG Effect* (to pass in the [de_bias.dat](#func.loess_de.args.de_bias.dat) argument to the internally called [loess_de](#func.loess_de) described above);

::: { #func.calc_UMGY_SmoothedDay_effect.args. .sidebar }
  - **degree:** (Integer, $NA$ by default) The degree of the polynomials to be used (passed in the same-named argument to the internally called [loess_de](#func.loess_de) described above);

:::
::: { #func.calc_UMGY_SmoothedDay_effect.args. .sidebar }
  - **span:** (Numeric, $NA$ by default ) The parameter $\alpha$, which controls the degree of smoothing (passed in the same-named argument to the internally called [loess_de](#func.loess_de) described above);

@@

This auxiliary function is a wrapper for the [loess_de](#func.loess_de) described above, intended to be used internally in the following helper functions (both described below in this section):

  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);

  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

An example of using it in the [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect) function is shown above in the [Usage](#func.calc_UMGY_SmoothedDay_effect.usage) section of *this Description*:
:::

@@

Below is a slightly simplified version of the source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) function:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect <- function(day_gen_effect, 
                                         degree = NA, 
                                         span = NA){
  fit <- day_gen_effect |> loess_de(degree, span)
  smth_day_effect <- day_gen_effect |> mutate(de_smoothed = fit$fitted)

  put_log2("Function `calc_UMGY_SmoothedDay_effect`: 
Model has been trained using `loess` function with the following parameters:
degree: %1;
span: %2.",
           degree,
           span)

  smth_day_effect
}
```

::: {.noteblock data-latex=""}
The complete source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) function is available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L3) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) script on _GitHub_.
:::
:::

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_SmoothedDay_effect.args.lambda) argument value is non-default) the *User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model* using the training sample of the `Movielens` dataset passed in the [train_set](#func.train_UMGY_SmoothedDay_effect.args.train_set) argument.

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_SmoothedDay_effect.cv.args.lambda) argument value is non-default) the *User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model* with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

Trains (and optionally *regularizes* if the [lambda](#func.regularize.train_UMGYD_effect.args.lambda) argument value is non-default) the *User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model* using the training sample of the `Movielens` dataset passed in the [train_set](#func.regularize.train_UMGYD_effect.args.train_set) argument.

@@

Trains and *regularizes* the *User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model* with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMGYD_effect.cv.args.lambda) argument for the *User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.

@@

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMGYD_effect.cv.details) section of *this Description* below.
:::
:::
